Yield Mapping

### 2016-02-10 Rmd/yield-predictors.rmd
Redid yield-predictors.Rmd with DSSAT-modeled maize, so greatly reduced now--no complicated merges.

# Predict DSSAT yields onto finer soil map

@@

p_root <- lmisc::set_base_path()
p_crop <- "~/Dropbox/data/agricultural/zambia/dssat/"
p_cropmod <- "~/Dropbox/projects/rCropMod/external/ext_data/"
p_soil <- "~/DSSAT45/SOIL/"
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_clim <- fp(p_root, "external/input_devel/climate")
p_climo <- "~/Dropbox/data/climate/rainfall/zambia/precip/"
p_yield <- fp(p_root, "external/input_devel/yield")
p_pgf <- "~/Dropbox/data/climate/afdm/pgf_bc/zambia/"
p_sgrids <- "~/Dropbox/data/soil/"

@@

# Read in soil-climate grid from Di's work
ID <- fread(full_path(p_crop, "Soil_clim_match_ID_master.csv"))
ID[, V1 := NULL]
setkeyv(ID, c("WTH", "SOL", "Zone"))
ID[, SharePct := as.numeric(SharePct)]
zone <- sort(unique(ID$Zone))
for(i in 1:length(zone)) ID[Zone == zone[i], zoneid := i]
setkeyv(ID, c("WTH", "Zone"))
# ID[, FID := 1:.N]
IDr <- subset(unique(ID[, .(x, y, ID, WTH, Zone, zoneid)]))
setkeyv(IDr, c("WTH", "Zone"))
# ID[SOL == "WI_ARBW438"]
gcs <- "+proj=longlat +datum=WGS84"
idgr <- dt_to_raster(IDr[, .(x, y, ID, zoneid)], CRS(gcs))
# plot(idgr)

# Yield data selection points
load(fp(p_yield, "empirical/stations.rda"))
zmetsamps <- extract(idgr, zmetf, cellnumbers = TRUE)
zmetxy <- data.frame(zmetf)[, c("x", "y")]
# xFromCell(idgr, zmetsamps[, "cells"]) - zmetxy$x
# yFromCell(idgr, zmetsamps[, "cells"]) - zmetxy$y
# zmetsamps <- cbind(xyFromCell(idgr, zmetsamps[, "cells"]), zmetsamps)
zmetsamps <- cbind(zmetxy, zmetsamps)  # harmonize with DSSAT zmetfxy

# zmet_dt <- data.table(data.frame(zmetsamps))[, .(x, y)]
zmet_dt <- data.table(data.frame(zmetsamps))
# setkeyv(zmet_dt, c("x", "y"))
setkeyv(zmet_dt, c("ID", "zoneid"))

# setkeyv(IDr, c("x", "y"))
setkeyv(IDr, c("ID", "zoneid"))
IDr_ss <- merge(IDr, zmet_dt)
# IDr_ss <- IDr[zmet_dt]  # pick up IDs, soils, wth variables
IDr_ss[, c("x.x", "y.x", "cells") := NULL]
setnames(IDr_ss, c("x.y", "y.y"), c("x", "y"))
setkeyv(IDr_ss, c("WTH", "Zone"))

# Collect soils up here also 
pts <- IDr_ss
# pts <- cbind.data.frame("x" = 27.4738917, "y" = -16.2783358)  # in Zambia
coordinates(pts) <- ~x+y
soilid <- extract(soilgrid, pts)  # grid id corresponding to point
IDr_ss[, soilid := soilid]
profs <- soilshare[CELL5M %in% soilid]  # profiles at that point
profs[, SharePct := NULL]
IDr_ss <- merge(IDr_ss, profs, by.x = "soilid", by.y = "CELL5M")
IDr_ss[, soilid := NULL]
setnames(IDr_ss, "SoilProfile", "SOL")
setkeyv(IDr_ss, c("Zone", "WTH", "SOL"))

@@

# Full DSSAT yield grids
```{r, eval = FALSE}
# crops <- dir(p_crop, pattern = "maize|soybean", full.names = FALSE)
crops <- dir(p_crop, pattern = "maize", full.names = FALSE)  # only maize now
fylds <- lapply(crops, function(X) {  # X <- "maize"
  fnms <- dir(full_path(p_crop, X), pattern = X)
  fl <- unlist(lapply(c("S", "R"), function(x) {
    fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
  }))  # reorder to read subsistence first
  # idt <- copy(IDr)
  # key(idt)
  idt <- copy(ID)
  idt <- subset(unique(idt[, .(x, y, WTH, Zone, SharePct)]))
  setkeyv(idt, c("WTH", "Zone"))
  for(y in 1:length(fl)) { # y <- 1
    # setcolorder(idt, c(4, 7:8, 5:6, 1:3))
    print(paste("Processing", fl[y]))
    cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
    cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
    setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
    cr[, WTH := substr(WTH, 1, 4)]
    cr[, c("V1", "RUNNO", "TRNO", "TNAM") := NULL]
    setkeyv(cr, c("WTH", "Zone"))

    crm <- merge(cr, idt, by = c("WTH", "Zone"))

    # nms <- names(crm)[c(6:7, 8, 1:2, 4, 9, 5, 3)]

    # Soil averaged yields
    solmu <- crm[, round(weighted.mean(HWAH, w = SharePct)), 
                 by = .(WTH, Zone, YR)]
    snm <- gsub("*.*_|\\.csv", "", fl[y])

    # Yield means
    v <- solmu[, round(mean(V1)), by = .(WTH, Zone)]
    setnames(v, "V1", paste0("Y", snm))
    idt <- idt[v]
    setkeyv(idt, c("WTH", "Zone"))
    idt
  }
  idt
})
names(fylds) <- crops

rdt <- fylds$maize[, c(1:2, grep("YS", names(fylds$maize))), with = FALSE]
mzs <- dt_to_raster(rdt, raster()@crs)
rdt <- fylds$maize[, c(1:2, grep("YR", names(fylds$maize))), with = FALSE]
mzr <- dt_to_raster(rdt, raster()@crs)

# par(mar = rep(0, 4))
plot(calc(mzs, max))
plot(calc(mzr, mean))
# plot(calc(soyr, mean))
# plot(zmetf, pch = 16, cex = 0.5, add = TRUE)

@@

# Extract subset of DSSAT yields, soils, climate data
## Yield
```{r, eval=FALSE}
# merge function for lists of data.tables
mymerge <- function(x, y) merge(x, y, all = TRUE)

crops <- dir(p_crop, pattern = "maize", full.names = FALSE)  # only maize now
# Yield data for those points
mannms <- c("S", "R")
ylds <- lapply(crops, function(X) {
  # X <- crops[1]
  fnms <- dir(full_path(p_crop, X), pattern = X)
  fl <- unlist(lapply(mannms, function(x) {
    fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
  }))  # reorder to read subsistence first
  # fl <- fl[1:3]
  crml <- lapply(1:length(fl), function(y) { # y <- 1
    idt <- copy(IDr_ss)
    print(paste("Processing", fl[y]))
    cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
    cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
    setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
    cr[, WTH := substr(WTH, 1, 4)]
    cr[, c("V1", "RUNNO", "TRNO", "TNAM") := NULL]
    setkeyv(cr, c("Zone", "WTH", "SOL"))

    # merge IDs with subset location
    # crm <- cr[idt]
    crm <- merge(cr, idt)
    # crm[, SharePct := NULL]
    nms <- names(crm)[c(7:8, 6, 1:2, 9, 3, 5, 4)]
    setcolorder(crm, nms)
    mid <- 1:nrow(crm)
    crm <- cbind(mid, crm)
    keys <- c("mid", "x", "y", "ID", "WTH", "Zone", "zoneid", "SOL", "YR")
    setkeyv(crm, keys)
    # setcolorder(crm, c(nms[grep("HWAH", nms, invert = TRUE)], "HWAH"))
    newnm <- gsub(".*._||\\.csv", "", fl[y])
    setnames(crm, "HWAH", newnm)
    crm
  })
  yldmerge <- Reduce(mymerge, crml)
  ynms <- names(yldmerge)
  # yldred <- data.table(round(do.call(cbind, lapply(mannms, function(x) {
    # yldmerge[, rowMeans(.SD), .SDcols = grep(paste0(x, "[0-9]"), ynms)]
  # }))))
  # calculate 
  for(i in mannms) {
    onms <- list(paste0("YLD", i), grep(paste0(i, "[0-9]"), ynms))
    # yldmerge[, c(onms[[1]]) := round(rowMeans(.SD)), .SDcols = onms[[2]]]
    # Selecting rowMax here
    yldmerge[, c(onms[[1]]) := apply(.SD, 1, max), .SDcols = onms[[2]]]
    yldmerge[, c(onms[[2]]) := NULL]
  }

  # Internal check on yield means
  if(length(crml) == 8) crind <- list("S" = 1:4, "R" = 5:8, "I" = 9:12) 
  if(length(crml) == 6) crind <- list("S" = 1:3, "R" = 4:6, "I" = 7:9) 
  for(ii in sample(1:nrow(yldmerge), size = 20)) {
    for(jj in 1:length(mannms)) {  # jj = 2
      a <- yldmerge[ii, ]
      b <- sapply(crind[[jj]], function(x) {
        crml[[x]][mid == a$mid, ][[ncol(crml[[x]])]]
      })
      #if(a[[paste0("YLD", mannms[jj])]] != round(mean(b))) {
      if(a[[paste0("YLD", mannms[jj])]] != round(max(b))) {
        print(paste("Yield mismatch at", ii, mannms[jj]))
      }
    }
  }

  # swap in averaged yields
  # yldmerge[, mid := NULL]
  # yldmerge <- cbind(yldmerge, yldred)
  yldmerge
})
names(ylds) <- crops

soy_ylds <- fread(fp(p_yield, "DSSAT/soy_ylds.csv"))
# soy_ylds[, year := as.numeric(substr(PDAT, 1, 4))]
soy_ylds[, YR := 1:31, by = .(XNAME, TRNO)]  # unique year
soy_ylds[, c("RUNNO", "PROF", "PDAT", "MDAT", "PRCM") := NULL]
setnames(soy_ylds, "SOIL_ID...", "SOL")

# calculate mean and max as YLDS and YLDR
soy_ylds2 <- soy_ylds[,  round(mean(HWAH)), by = .(XNAME, x, y, WTH, SOL, YR)]
setnames(soy_ylds2, "V1", "YLDS")
soy_ylds2[, YLDR := soy_ylds[,  round(max(HWAH)), by = .(XNAME, YR)][, V1]]
soy_ylds2[, XNAME := NULL]

ylds$maize[, c("ID", "Zone", "zoneid") := NULL]
soy_ylds2 <- cbind(c("mid" = 1:nrow(soy_ylds2)), soy_ylds2)
setnames(soy_ylds2, "V1", "mid")

ylds <- list(ylds$maize, soy_ylds2)
names(ylds) <- c("maize", "soybean")

@@

## Soil
### Profile data
Note: Using an updated version of `read_dssat_prof.R` from the rCropMod library. It needs to be rebuilt into that library.  Here it is being sourced before runnign the next section. 
```{r, eval = FALSE}
pts <- IDr_ss[, .(x, y)]
# pts <- cbind.data.frame("x" = 27.4738917, "y" = -16.2783358)  # in Zambia
coordinates(pts) <- ~x+y
soilid <- extract(soilgrid, pts)  # grid id corresponding to point
profs <- soilshare[CELL5M %in% soilid]  # profiles at that point
profdat <- read_fullSOL_prof(solfile = full_path(p_soil, "WI.SOL"), 
                             profiles = profs$SoilProfile) 
solvars <- unique(gsub("[[:digit:]]", "", names(profdat)))
lnms <- names(profdat)

@@

# finalize
sol1m <- cbind("prof" = profdat[, prof], data.table(solvals))
# unique(ylds$soybean$SOL) %in% sol1m$prof
```

@@

# variabes to derive: MAP, N raindays, rainfall depth, n rain days, 
# n dry days in Jan and Feb, mean max temp, N days > 30, annual mean temp, 
# GDD > 30, average daily SRAD
# set up indices for selecting dates
ystrt <- which(mos == 7 & dys == 15) 
ystrt <- ystrt[-length(ystrt)]
yend <- which(mos == 7 & dys == 14)[-1]
# set up seasonal index
sind <- rep(0, length(dts))  # whole year spanning growing season
for(i in 1:length(ystrt)) sind[ystrt[i]:yend[i]] <- i
sindu <- unique(sind)[-1]
janind <- rep(0, length(dts))  # january of each year
for(i in sindu) janind[which(sind == i & mos == 1)] <- i
febind <- rep(0, length(dts))  # feb of each year
for(i in sindu) febind[which(sind == i & mos == 2)] <- i
jfmind <- rep(0, length(dts))  # djf of each growing season
for(i in sindu) jfmind[which(sind == i & mos %in% c(1, 2, 3))] <- i
ndjfind <- rep(0, length(dts))  # djf of each growing season
for(i in sindu) ndjfind[which(sind == i & mos %in% c(1, 2, 11, 12))] <- i
octind <- rep(0, length(dts))  # feb of each year
for(i in sindu) octind[which(sind == i & mos == 10)] <- i

@@

# storm depth in NDJF
# sdeps <- lapply(sindu2, function(x) { # x <- sindu2[1]
#   print(x)
#   ryr <- prec[[which(ndjfind == x)]]
#   r <- calc(ryr, sdepth)
#   names(r) <- as.character(ano[x])
#   r
# })
# sdeps <- stack(sdeps)
# 
# # storm depth in NDJF
# rdays <- lapply(sindu2, function(x) { # x <- sindu2[1]
#   print(x)
#   ryr <- prec[[which(djfind == x)]]
#   r <- calc(ryr, nrday)
#   names(r) <- as.character(ano[x])
#   r
# })
# rdays <- stack(rdays)
# 
# # dry days in January
# drydaysj <- lapply(sindu2, function(x) { # x <- sindu2[1]
#   print(x)
#   ryr <- prec[[which(janind == x)]]
#   r <- calc(ryr, ndry)
#   names(r) <- as.character(ano[x])
#   r
# })
# drydaysj <- stack(drydaysj)
# 
# # dry days in February
# drydaysf <- lapply(sindu2, function(x) { # x <- sindu2[1]
#   print(x)
#   ryr <- prec[[which(febind == x)]]
#   r <- calc(ryr, ndry)
#   names(r) <- as.character(ano[x])
#   r
# })
# drydaysf <- stack(drydaysf)

@@

@@

# par(mfrow = c(6, 6), mar = c(0, 0, 0, 0))
# for(i in 1:nlayers(octtot)) {
#   plot(octtot[[i]], axes = FALSE, box = FALSE)
#   plot(prov, add = TRUE, zlim = c(0, 1000))
# }
# which(1980:2010 %in% ) 
# pre2005 <- calc(octtot[[1:26]], mean)
# post2005 <- calc(octtot[[27:31]], mean)
# 
# par(mfrow = c(1, 2), mar = c(0, 0, 0, 0))
# plot(pre2005, axes = FALSE, box = FALSE, zlim = c(0, 400))
# plot(prov, add = TRUE)
# plot(post2005, axes = FALSE, box = FALSE, zlim = c(0, 400))
# plot(prov, add = TRUE)
# 
# par(mfrow = c(1, 1), mar = c(0, 0, 0, 0))
# plot((pre2005 - post2005) / pre2005 * 100, zlim = c(-100, 100))
# plot(prov, add = TRUE)

@@

# Create tmean variable
tmu <- foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(sind == sindu2[i])
  tmu <- (tmax[[ind]] + tmin[[ind]]) / 2 - 273.15
}

# Functions
# GDD between 8-30 degrees
gdd830 <- function(x) {
  a <- x[x < 30] - 8
  sum(ifelse(a < 0, 0, a))
}
# x <- sample(4:40, size = 100, replace = T)
# gdd830(x)

@@

# growing degree day and killing degree days
kdd <- stack(kdy <- foreach(i = 1:length(tmu)) %dopar% calc(tmu[[i]], kdd))
writeRaster(kdd, filename = fp(p_clim, "kdd.tif"), overwrite = TRUE)
gdd <- stack(gdy <- foreach(i = 1:length(tmu)) %dopar% calc(tmu[[i]], gdd830))
writeRaster(gdd, filename = fp(p_clim, "gdd.tif"), overwrite = TRUE)

@@

1. Predictors: 
    + Weather/climate: 
        +`rftot`, `ndjftot`, `jantot`, `febtot` for rainfall; `kdays`, `gdd`, `f30jan`, `f30feb`, `tmax_gs`, `tmax_ano` for temperature; `srtot_gs`, `srmu_gs`, `srtot_ano`, and `srmu_ano` for solar radiation
    + Soil: mean soil properties within top 1 m
    + Management variables: Codes S, R, I
2. Predictands: mean yields across cultivars within each management type

>>>>

# Develop predictors for two-stage yield model

@@

setwd("~/Dropbox/publications/zambiaToff/private/zambiaToff/")
p_root <- lmisc::set_base_path()
p_crop <- "~/Dropbox/data/agricultural/zambia/dssat/"
p_cropmod <- "~/Dropbox/projects/rCropMod/external/ext_data/"
p_soil <- "~/DSSAT45/SOIL/"
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_clim <- fp(p_root, "external/input_devel/climate")
p_climo <- "~/Dropbox/data/climate/rainfall/zambia/precip/"
p_yield <- fp(p_root, "external/input_devel/yield")
p_pgf <- "~/Dropbox/data/climate/afdm/pgf_bc/zambia/"
p_sgrids <- "~/Dropbox/data/soil/"
p_dssat <- fp(p_yield, "DSSAT")

@@

# Read in soil-climate grid from Di's work
# ID <- fread(full_path(p_crop, "Soil_clim_match_ID_master.csv"))
# ID[, V1 := NULL]
# setkeyv(ID, c("WTH", "SOL", "Zone"))
# ID[, SharePct := as.numeric(SharePct)]
# zone <- sort(unique(ID$Zone))
# for(i in 1:length(zone)) ID[Zone == zone[i], zoneid := i]
# setkeyv(ID, c("WTH", "Zone"))
# # ID[, FID := 1:.N]
# IDr <- subset(unique(ID[, .(x, y, ID, WTH, Zone, zoneid)]))
# setkeyv(IDr, c("WTH", "Zone"))
# # ID[SOL == "WI_ARBW438"]
# idgr <- dt_to_raster(IDr[, .(x, y, ID, zoneid)], CRS(gcs))
# plot(idgr)

# Yield data selection points
gcs <- "+proj=longlat +datum=WGS84"
load(fp(p_yield, "empirical/stations.rda"))
zmetf@proj4string <- CRS(gcs)

# zmetsamps <- extract(idgr, zmetf, cellnumbers = TRUE)
# zmetxy <- data.frame(zmetf)[, c("x", "y")]
zmetxy <- data.table(data.frame(zmetf)[, 1:3])
wthnms <- substr(zmetxy$nms, 1, 4)
ginds <- grep("g", wthnms)
combis <- expand.grid(data.frame(cbind(LETTERS, LETTERS, LETTERS)))[, 3:1]
combis <- paste0(combis[, 1], combis[, 2], combis[, 3])
wthnms[ginds] <- paste0("G", combis[1:length(ginds)])
zmetxy <- cbind(data.table("WTH" = wthnms), zmetxy)

# xFromCell(idgr, zmetsamps[, "cells"]) - zmetxy$x
# yFromCell(idgr, zmetsamps[, "cells"]) - zmetxy$y
# zmetsamps <- cbind(xyFromCell(idgr, zmetsamps[, "cells"]), zmetsamps)
# zmetsamps <- cbind(zmetxy, zmetsamps)  # harmonize with DSSAT zmetfxy

# zmet_dt <- data.table(data.frame(zmetsamps))[, .(x, y)]
# zmet_dt <- data.table(data.frame(zmetsamps))
# # setkeyv(zmet_dt, c("x", "y"))
# setkeyv(zmet_dt, c("ID", "zoneid"))
# 
# # setkeyv(IDr, c("x", "y"))
# setkeyv(IDr, c("ID", "zoneid"))
# IDr_ss <- merge(IDr, zmet_dt)
# # IDr_ss <- IDr[zmet_dt]  # pick up IDs, soils, wth variables
# IDr_ss[, c("x.x", "y.x", "cells") := NULL]
# setnames(IDr_ss, c("x.y", "y.y"), c("x", "y"))
# setkeyv(IDr_ss, c("WTH", "Zone"))

# Collect soils up here also 
# pts <- IDr_ss
# pts <- cbind.data.frame("x" = 27.4738917, "y" = -16.2783358)  # in Zambia
# coordinates(pts) <- ~x+y
# soilid <- extract(soilgrid, pts)  # grid id corresponding to point
# soilid <- extract(soilgrid, zmetf)  # grid id corresponding to point
# IDr_ss[, soilid := soilid]
# profs <- soilshare[CELL5M %in% soilid]  # profiles at that point
# profs[, SharePct := NULL]
# IDr_ss <- merge(IDr_ss, profs, by.x = "soilid", by.y = "CELL5M")
# IDr_ss[, soilid := NULL]
# setnames(IDr_ss, "SoilProfile", "SOL")
# setkeyv(IDr_ss, c("Zone", "WTH", "SOL"))

@@

# Full DSSAT yield grids
```{r, eval = FALSE}
# # crops <- dir(p_crop, pattern = "maize|soybean", full.names = FALSE)
# crops <- dir(p_crop, pattern = "maize", full.names = FALSE)  # only maize now
# fylds <- lapply(crops, function(X) {  # X <- "maize"
#   fnms <- dir(full_path(p_crop, X), pattern = X)
#   fl <- unlist(lapply(c("S", "R"), function(x) {
#     fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
#   }))  # reorder to read subsistence first
#   # idt <- copy(IDr)
#   # key(idt)
#   idt <- copy(ID)
#   idt <- subset(unique(idt[, .(x, y, WTH, Zone, SharePct)]))
#   setkeyv(idt, c("WTH", "Zone"))
#   for(y in 1:length(fl)) { # y <- 1
#     # setcolorder(idt, c(4, 7:8, 5:6, 1:3))
#     print(paste("Processing", fl[y]))
#     cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
#     cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
#     setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
#     cr[, WTH := substr(WTH, 1, 4)]
#     cr[, c("V1", "RUNNO", "TRNO", "TNAM") := NULL]
#     setkeyv(cr, c("WTH", "Zone"))
#     
#     crm <- merge(cr, idt, by = c("WTH", "Zone"))
# 
#     # nms <- names(crm)[c(6:7, 8, 1:2, 4, 9, 5, 3)]
#     
#     # Soil averaged yields
#     solmu <- crm[, round(weighted.mean(HWAH, w = SharePct)), 
#                  by = .(WTH, Zone, YR)]
#     snm <- gsub("*.*_|\\.csv", "", fl[y])
#     
#     # Yield means
#     v <- solmu[, round(mean(V1)), by = .(WTH, Zone)]
#     setnames(v, "V1", paste0("Y", snm))
#     idt <- idt[v]
#     setkeyv(idt, c("WTH", "Zone"))
#     idt
#   }
#   idt
# })
# names(fylds) <- crops
# 
# rdt <- fylds$maize[, c(1:2, grep("YS", names(fylds$maize))), with = FALSE]
# mzs <- dt_to_raster(rdt, raster()@crs)
# rdt <- fylds$maize[, c(1:2, grep("YR", names(fylds$maize))), with = FALSE]
# mzr <- dt_to_raster(rdt, raster()@crs)
# 
# # par(mar = rep(0, 4))
# plot(calc(mzs, max))
# plot(calc(mzr, mean))
# # plot(calc(soyr, mean))
# # plot(zmetf, pch = 16, cex = 0.5, add = TRUE)

@@

# Extract subset of DSSAT yields, soils, climate data
## Yield
```{r, eval=FALSE}
# mymerge <- function(x, y) merge(x, y, all = TRUE)
crops <- c("maize", "soy")
ylds <- lapply(crops, function(X) {  # X <- crops[2]
  fnms <- dir(p_dssat, pattern = paste0(X, "_"), full.names = TRUE)
  cr <- fread(fnms, stringsAsFactors = FALSE)
  cr[, YR := RUNNO - min(RUNNO) + 1, by = .(XNAME, TRNO)]  # unique year  
  cr[, c("SOIL_ID...", "RUNNO") := NULL]

  # Yields. S becomes mean across cultivars, R becomes max of cultivars
  cr2 <- cr[, round(mean(HWAH)), by = .(XNAME, x, y, WTH, PROF, YR)]
  setnames(cr2, c("V1", "PROF"), c("YLDS", "SOL"))
  cr2[, YLDR := cr[, round(max(HWAH)), by = .(XNAME, YR)][, V1]]
  cr2
})
names(ylds) <- crops

# crops <- dir(p_crop, pattern = "maize", full.names = FALSE)  # only maize now
# # Yield data for those points
# mannms <- c("S", "R")
# ylds <- lapply(crops, function(X) {
#   # X <- crops[1]
#   fnms <- dir(full_path(p_crop, X), pattern = X)
#   fl <- unlist(lapply(mannms, function(x) {
#     fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
#   }))  # reorder to read subsistence first
#   # fl <- fl[1:3]
#   crml <- lapply(1:length(fl), function(y) { # y <- 1
#     idt <- copy(IDr_ss)
#     print(paste("Processing", fl[y]))
#     cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
#     cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
#     setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
#     cr[, WTH := substr(WTH, 1, 4)]
#     cr[, c("V1", "RUNNO", "TRNO", "TNAM") := NULL]
#     setkeyv(cr, c("Zone", "WTH", "SOL"))
# 
#     # merge IDs with subset location
#     # crm <- cr[idt]
#     crm <- merge(cr, idt)
#     # crm[, SharePct := NULL]
#     nms <- names(crm)[c(7:8, 6, 1:2, 9, 3, 5, 4)]
#     setcolorder(crm, nms)
#     mid <- 1:nrow(crm)
#     crm <- cbind(mid, crm)
#     keys <- c("mid", "x", "y", "ID", "WTH", "Zone", "zoneid", "SOL", "YR")
#     setkeyv(crm, keys)
#     # setcolorder(crm, c(nms[grep("HWAH", nms, invert = TRUE)], "HWAH"))
#     newnm <- gsub(".*._||\\.csv", "", fl[y])
#     setnames(crm, "HWAH", newnm)
#     crm
#   })
#   yldmerge <- Reduce(mymerge, crml)
#   ynms <- names(yldmerge)
#   # yldred <- data.table(round(do.call(cbind, lapply(mannms, function(x) {
#     # yldmerge[, rowMeans(.SD), .SDcols = grep(paste0(x, "[0-9]"), ynms)]
#   # }))))
#   # calculate 
#   for(i in mannms) {
#     onms <- list(paste0("YLD", i), grep(paste0(i, "[0-9]"), ynms))
#     # yldmerge[, c(onms[[1]]) := round(rowMeans(.SD)), .SDcols = onms[[2]]]
#     # Selecting rowMax here
#     yldmerge[, c(onms[[1]]) := apply(.SD, 1, max), .SDcols = onms[[2]]]
#     yldmerge[, c(onms[[2]]) := NULL]
#   }
#   
#   # Internal check on yield means
#   if(length(crml) == 8) crind <- list("S" = 1:4, "R" = 5:8, "I" = 9:12) 
#   if(length(crml) == 6) crind <- list("S" = 1:3, "R" = 4:6, "I" = 7:9) 
#   for(ii in sample(1:nrow(yldmerge), size = 20)) {
#     for(jj in 1:length(mannms)) {  # jj = 2
#       a <- yldmerge[ii, ]
#       b <- sapply(crind[[jj]], function(x) {
#         crml[[x]][mid == a$mid, ][[ncol(crml[[x]])]]
#       })
#       #if(a[[paste0("YLD", mannms[jj])]] != round(mean(b))) {
#       if(a[[paste0("YLD", mannms[jj])]] != round(max(b))) {
#         print(paste("Yield mismatch at", ii, mannms[jj]))
#       }
#     }
#   }
#   
#   # swap in averaged yields
#   # yldmerge[, mid := NULL]
#   # yldmerge <- cbind(yldmerge, yldred)
#   yldmerge
# })
# names(ylds) <- crops
# 
# soy_ylds <- fread(fp(p_yield, "DSSAT/soy_ylds.csv"))
# # soy_ylds[, year := as.numeric(substr(PDAT, 1, 4))]
# soy_ylds[, YR := 1:31, by = .(XNAME, TRNO)]  # unique year
# soy_ylds[, c("RUNNO", "PROF", "PDAT", "MDAT", "PRCM") := NULL]
# setnames(soy_ylds, "SOIL_ID...", "SOL")
# 
# # calculate mean and max as YLDS and YLDR
# soy_ylds2 <- soy_ylds[,  round(mean(HWAH)), by = .(XNAME, x, y, WTH, SOL, YR)]
# setnames(soy_ylds2, "V1", "YLDS")
# soy_ylds2[, YLDR := soy_ylds[,  round(max(HWAH)), by = .(XNAME, YR)][, V1]]
# soy_ylds2[, XNAME := NULL]
# 
# ylds$maize[, c("ID", "Zone", "zoneid") := NULL]
# soy_ylds2 <- cbind(c("mid" = 1:nrow(soy_ylds2)), soy_ylds2)
# setnames(soy_ylds2, "V1", "mid")
# 
# ylds <- list(ylds$maize, soy_ylds2)
# names(ylds) <- c("maize", "soybean")

@@

## Soil
### Profile data
Note: Using an updated version of `read_dssat_prof.R` from the rCropMod library. It needs to be rebuilt into that library.  Here it is being sourced before runnign the next section. 
```{r, eval = FALSE}
# pts <- IDr_ss[, .(x, y)]
# pts <- cbind.data.frame("x" = 27.4738917, "y" = -16.2783358)  # in Zambia
# coordinates(pts) <- ~x+y
soilid <- extract(soilgrid, zmetf)  # grid id corresponding to point
# soilid <- extract(soilgrid, pts)  # grid id corresponding to point
profs <- soilshare[CELL5M %in% soilid]  # profiles at that point
profdat <- read_fullSOL_prof(solfile = full_path(p_soil, "WI.SOL"), 
                             profiles = profs$SoilProfile)
setnames(profdat, "prof", "SOL")
solvars <- unique(gsub("[[:digit:]]", "", names(profdat)))
lnms <- names(profdat)

@@

# finalize
sol1m <- cbind("SOL" = profdat[, SOL], data.table(solvals))
# unique(ylds$soybean$SOL) %in% sol1m$prof
```

@@

# variabes to derive: MAP, N raindays, rainfall depth, n rain days, 
# n dry days in Jan and Feb, mean max temp, N days > 30, annual mean temp, 
# GDD > 30, average daily SRAD
# set up indices for selecting dates
ystrt <- which(mos == 7 & dys == 15) 
ystrt <- ystrt[-length(ystrt)]
yend <- which(mos == 7 & dys == 14)[-1]
# set up seasonal index
sind <- rep(0, length(dts))  # whole year spanning growing season
for(i in 1:length(ystrt)) sind[ystrt[i]:yend[i]] <- i
sindu <- unique(sind)[-1]
janind <- rep(0, length(dts))  # january of each year
for(i in sindu) janind[which(sind == i & mos == 1)] <- i
febind <- rep(0, length(dts))  # feb of each year
for(i in sindu) febind[which(sind == i & mos == 2)] <- i
jfmind <- rep(0, length(dts))  # djf of each growing season
for(i in sindu) jfmind[which(sind == i & mos %in% c(1, 2, 3))] <- i
ndjfind <- rep(0, length(dts))  # djf of each growing season
for(i in sindu) ndjfind[which(sind == i & mos %in% c(1, 2, 11, 12))] <- i
octind <- rep(0, length(dts))  # feb of each year
for(i in sindu) octind[which(sind == i & mos == 10)] <- i
djfmind <- rep(0, length(dts))  # feb of each year
for(i in sindu) djfmind[which(sind == i & mos  %in% c(12, 1:3))] <- i
# cbind(yrs[sind == 1], mos[sind == 1], djfmind[sind == 1])

@@

# storm depth in NDJF
# sdeps <- lapply(sindu2, function(x) { # x <- sindu2[1]
#   print(x)
#   ryr <- prec[[which(ndjfind == x)]]
#   r <- calc(ryr, sdepth)
#   names(r) <- as.character(ano[x])
#   r
# })
# sdeps <- stack(sdeps)
#

@@

# DJFM season rainfall
djfmtot <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(djfmind == sindu2[i])
  calc(prec[[ind]], sum)
})
writeRaster(djfmtot, filename = fp(p_clim, "djfmtot.tif"), overwrite = TRUE)

@@

@@

# Create tmean variable
tmu <- foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(sind == sindu2[i])
  tmu <- (tmax[[ind]] + tmin[[ind]]) / 2 - 273.15
}
writeRaster(stack(tmu), filename = fp(p_clim, "tmu.tif"), overwrite = TRUE)

# Functions
gddf <- function(x) {
  a <- x - 8
  sum(ifelse(a < 0, 0, a))
}

# GDD between 8-30 degrees
gdd830 <- function(x) {
  a <- x[x < 30] - 8
  sum(ifelse(a < 0, 0, a))
}
# x <- sample(4:40, size = 100, replace = T)
# gdd830(x)

@@

# growing degree day and killing degree days
kdd <- stack(kdy <- foreach(i = 1:length(tmu)) %dopar% calc(tmu[[i]], kdd))
writeRaster(kdd, filename = fp(p_clim, "kdd.tif"), overwrite = TRUE)
gdd30 <- stack(gdy <- foreach(i = 1:length(tmu)) %dopar% calc(tmu[[i]], gdd830))
writeRaster(gdd30, filename = fp(p_clim, "gdd30.tif"), overwrite = TRUE)
gdd <- stack(gdy <- foreach(i = 1:length(tmu)) %dopar% calc(tmu[[i]], gddf))
writeRaster(gdd30, filename = fp(p_clim, "gdd.tif"), overwrite = TRUE)

@@

1. Predictors: 
    + Weather/climate: 
        +`rftot`, `ndjftot`, `jantot`, `febtot` for rainfall; `kdays`, `gdd`, `f30jan`, `f30feb`, `tmax_gs`, `tmax_ano` for temperature; `srtot_gs`, `srmu_gs`, `srtot_ano`, and `srmu_ano` for solar radiation (plus others added on the way)
    + Soil: mean soil properties within top 1 m
    + Management parameters: Codes S, R, 
2. Predictands: mean yields across cultivars within each management type
